\documentclass{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage[shortlabels]{enumitem}

\title{1.2 Distribution and Factoring Homework Set}
\author{}
\date{}

\begin{document}
\maketitle

\subsection*{Distribution and Factoring Problems}
\begin{enumerate}
    \item If $ab = 7$, $bc = 5$ and $a + c = 4$, what is $a + b + c$?
    \vspace{3cm}
    \item Evaluate $100002^2 - 99998^2$.
    \vspace{3cm}
    \item The sum of two numbers is $11$ and their product is $13$. What
    is twice the sum of their reciprocals? Express your answer as a
    common fraction.
    \vspace{3cm}
    \item The difference between the squares of two numbers is $80$. If the
    sum of the two numbers is $16$, what is their positive difference?
    \vspace{3cm}
    \item The sum of two positive numbers is $6$, and the sum of their
    squares is $22$. What is the sum of their reciprocals? Express your
    answer as a common fraction.
\end{enumerate}

\newpage

\subsection*{Systems of Equations Review Problems}
\begin{enumerate}[resume]
    \item Four neighbors go to purchase letters to spell their house numbers
    at a local hardware store. Each letter is priced separately. The first
    neighbor, who lives in house number one, purchases the letters ONE for
    $\$2$. The second neighbor buys TWO for $\$3$ and the third pays $\$5$
    for the letters to spell ELEVEN. How much will he letters in the word
    TWELVE cost thr fourth neighbor?
    \vspace{3cm}
    \item The units digit in a two-digit number is three times the tens
    digit. If the digits are reversed, the resulting number is $54$ more
    than the original number. Find the original number.
    \vspace{3cm}
    \item George has a farm with pigs and chickens. He counted $27$ heads
    and $94$ legs. Assuming all the animals on the farm have all of their
    limbs, how many chickens are there?
    \vspace{3cm}
\end{enumerate}
\end{document}
