\documentclass{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage[shortlabels]{enumitem}
\usepackage{amsmath}

\title{1.3 Arithmetic Homework}
\author{}
\date{}

\begin{document}
\maketitle

\subsection*{Arithmetic}
\begin{enumerate}
    \item To make this statement true, what operator should replace the question mark between the $6$ and the $3$?
    Examples of math operators include $\div$, $\times$, $+$, and $-$.
    \[(6?3) + 4 - (2 - 1) = 5\] 
    \vspace{3cm}
    \item If $a@b = \frac{a \times b}{a + b}$ for $a, b$ positive integers, then what is $5@10$?
    \vspace{3cm}
    \item How many digits are in the product $4^{5} \cdot 5^{10}$?
    \vspace{3cm} 
\end{enumerate}

\subsection*{Systems of Equations Review Problems}
\begin{enumerate}[resume]
    \item Four neighbors go to purchase letters to spell their house numbers at a local hardware store.
    Each letter is priced separately.
    The first neighbor, who lives in house number one, purchases the letters ONE for $\$2$.
    The second neighbor buys TWO for $\$3$ and the third pays $\$5$ for the letters to spell ELEVEN.
    How much will the letters in the word TWELVE cost the fourth neighbor?
    \vspace{3cm}
    \item The units digit in a two-digit number is three times the tens digit.
    If the digits are reversed, the resulting number is $54$ more than the original number.
    Find the original number.
    \vspace{3cm}
    \item George has a farm with pigs and chickens.
    He counted $27$ heads and $94$ legs.
    Assuming all the animals on the farm have all of their limbs, how many chickens are there?
    \vspace{3cm}
\end{enumerate}

\end{document}
