\documentclass{article}

\usepackage{enumitem}
\usepackage{siunitx}
\usepackage{amsmath}

\title{1.1 Systems of Equations Notes}
\author{}
\date{}

\begin{document}
    \maketitle
    \section*{Important Concepts}
    \subsection*{Variables, Expressions, and Equations}
    A \textbf{variable} is a letter used to represent a potentially unknown value.
    $x$ is often used by convention to denote a value in an equation which we're trying to find.
    An \textbf{expression} is a combination of \emph{symbols}, which can include numbers such as $42$, operators such as $+$, and variables such as $x$.
    An \textbf{equation} is a statement which states that two expressions are equal.
    For example, $2x - 7 = x - 8$ is an equation which states that the values of the expression $2x - 7$ and the expression $x - 8$ are equal.
    \subsection*{Notation}
    In algebra and more advanced mathematics, the $\times$ sign is rarely used to denote multiplication.
    Instead, the $\cdot$ sign is used.
    For example, we write $3 \cdot 4$ instead of $3 \times 4$.
    The multiplication operator may also be completely omitted if it doesn't cause ambiguity.
    For example, $ab$ means $a$ times $b$, and $3x$ means $3$ times $x$.
    When we multiply a number by a variable, we typically write the number first, so we would write $4x$ instead of $x4$.
    We also rarely use the $\div$ sign.
    Instead, we use the $/$ sign or the fraction bar.
    \subsection*{Substituting Values}
    The value of $2x - 7$ depends on the value of $x$.
    If we know the value of $x$, we can \textbf{substitute} it into the expression to find the value of the expression.
    For example, if $x = 4$, then we can replace the $x$ in $2x - 7$ with $4$, so $2x - 7 = 2 \cdot 4 - 7 = 8 - 7 = 1$.
    \subsection*{Solving Equations}
    When we have an equation involving variables, it's often useful to determine what values can be assigned to the variables such that the equation is true.
    We say that a value for an variable \textbf{satisfies} an equation if substituting that value into the equation results in a true statement, and we call such a value a \textbf{solution}.

    If we know that two expressions are equal, we can apply the same operation to both sides and they will still be equal.
    This is an important fact that we can use to solve equations.
    For example, if we know that $x + 2 = 5$, then we know that subtracting $2$ from both sides will result in two equal values.
    Therefore $x + 2 - 2 = 5 - 2$, so $x = 3$.
    By applying the same operation to both sides, we can \emph{isolate} the variable so that we can determine its value.
    For more complicated equations, we might need to apply multiple operations as shown in the video.
    If a variable appears multiple times, we will need to ``combine'' them by writing addition as multiplication or factoring.
    For example, $2x + 3x = (2 + 3)x = 5x$.
    \subsection*{Systems of Equations}
    An equation with two variables generally has infinitely many solutions.
    For example, some of the solutions of $y = 3x + 5$ are $(x, y) = (0, 5)$, $(x, y) = (1, 8)$, and $(x, y) = (2, 11)$.
    If we add in a second equation involving the two variables, we get a \textbf{system of equations} which usually have one solution that satisfies both equations.
    For example, the single solution to $y = 3x + 5$ and $2y = x + 9$ is $x = -\frac{1}{5}, y = \frac{22}{5}$.
    To find the solutions to systems of equations, we can use \textbf{substitution} and \textbf{elimination}.
    These methods transform a system of two equations into a single equation involving one variable.
    \subsubsection*{Substitution}
    To solve a system of equations by substitution, we isolate one of the variables in one equation and substitute the expression into the other equations.
    For example, to solve $y = 3x + 5$ and $2y = x + 9$, we notice that in the first equation, $y$ is already isolated.
    Therefore, we can substitute the expression which $y$ is equal to into the second equation to get $2(3x + 5) = x + 9$.
    This can then be solved using the methods for single variable equations to find the value of $x$.
    Once we know $x$, we can substitute its value into the first equation to find $y$.
    \subsubsection*{Elimination}
    We know that adding the same value to both sides of an equation results in an equivalent equation.
    If we know that $x + y = 8$ and $x - y = 4$, we can add $x - y$ to the left side of the first equation and $4$ to the right side since they are equal.
    Therefore, $x + y + x - y = 8 + 4$, so $2x = 12$ and we've eliminated one of the variables.
    We can then solve for $x$ and use the value of $x$ to find $y$.
    Sometimes, we may have to multiply the equations by some constants before adding or subtracting them so that one of the variables is eliminated.
    For example, if we have $x + y = 42$ and $2x + 3y = 2$, we can first multiply the first equation by $-2$ to get $-2x - 2y = -84$ and then add the two equations to eliminate $2x$.

    \section*{Example Problem}
    You can model many word problems with a system of equations.
    In this case, we already have a system set up for you.
    \begin{align}
        x + y &= 6 \label{eq1} \\
        x - y &= 5 \label{eq2}
    \end{align}
    We have a choice of substitution and elimination.
    We'll first try substitution, and then we'll cover elimination.
    
    \subsection*{Substitution}
    \begin{enumerate}
    \item Add $y$ to the left and right side of the equal sign of equation \eqref{eq2} to isolate the $x$ variable to one side.
    \begin{align}
        x - y &= 5 \nonumber \\
        x &= 5 + y \label{eq3}
    \end{align} 
    \item Then, substitute this new equation \eqref{eq3} into equation \eqref{eq1} and simplify.
    \begin{align*}
        x + y &= 6 \\
        (5 + y) + y &= 6 \\
        2y + 5 &= 6 \\
        2y &= 1 \\
        y &= \frac{1}{2}
    \end{align*}
    \item After solving for $y$, we can substitute this equation back in to equation \eqref{eq1} to solve for $x$.
    \begin{align*}
        x + y &= 6 \\
        x + \frac{1}{2} &= 6 \\
        x &= \frac{11}{2}
    \end{align*} 
    \item We can check our answers by plugging in these $x$ and $y$ values into equation \eqref{eq2}.
    \begin{align*}
        x + y &= 6 \\
        \frac{11}{2} + \frac{1}{2} &= 6
    \end{align*}
    \end{enumerate}

    \subsection*{Elimination}
    \begin{enumerate}
        \item Add equations \eqref{eq1} and \eqref{eq2} together and simplify.
        
        Notice that we are left with a single-variable equation.
        \begin{align*}
            x + y &= 6 \nonumber \\
            x - y &= 5 \nonumber \\
            (x + y) + (x - y) &= 6 + 5 \\
            2x &= 11 \\
            x &= \frac{11}{2}
        \end{align*}
        \item Substitute the $x$ value into equation \eqref{eq1} to solve for $y$.
        \begin{align*}
            x + y &= 6 \\
            \frac{11}{2} + y &= 6 \\
            y &= \frac{1}{2}
        \end{align*}
        \item Check your work by plugging in the $x$ and $y$ values into another equation.
    \end{enumerate}
\end{document}
