\documentclass{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{enumitem}
\usepackage[group-separator={,}]{siunitx}

\title{2.1 Statistics and Probability Lecture Problems}
\author{}
\date{}

\begin{document}
\maketitle

\subsection*{Statistics}
\begin{enumerate}
	\item Find the mode, median, arithmetic mean, range, and geometric mean
		of the following numbers: $2, 4, 4, 5, 6, 10$.
		\vspace{3cm}
	\item A student has an average score of $80$ on her first four tests.
		What must she score on the next test to raise her average to $82$?
		\vspace{3cm}
	\item The arithmetic mean of $12$ scores is $82$. When the highest and
		lowest scores are removed, the new mean becomes $84$. If the highest
		of the $12$ scores is $98$, what is the lowest score?
		\vspace{3cm}
	\item What is the average of seven numbers if the average of the first
		two is $9$ and the average of the last $5$ is $16$?
		\vspace{3cm}
\end{enumerate}

\newpage

\subsection*{Probability}
\begin{enumerate}[resume]
	\item A box contains three red balls and three green balls. Two balls
		are chosen.
		\begin{enumerate}
			\item What is the probability that a red ball is chosen first,
				then a green ball?
				\vspace{3cm}
			\item What is the probability that one red ball and one green
				ball are chosen?
				\vspace{3cm}
		\end{enumerate}
	\item John and Jayne each choose an integer from $1$ to $10$ inclusive.
		What is the probability that they each pick a number greater than
		$7$?
		\vspace{3cm}
	\item If Saphira randomly chooses a $4$-digit number, what is the
		probability that all four digits will be distinct?
\end{enumerate}
\end{document}